\documentclass[11pt]{article}

%%TO EDIT
\newcommand{\dueclassnumber}{9}
\newcommand{\assignmentnum}{4}

% CHANGE issolution{0} to issolution{1} for homework submission.
% WRITE solutions inside the \solution{} commands.
% or, you can use \if\issolution1 … \fi

\def\issolution{1}
\def\myname{Allen Liu} % My name goes here
\def\mysec{01} % Section number goes here
\def\myCM{374} % Campus Mailbox goes here

\input{../common/flags}
\newcommand\vv[2]{\begin{bmatrix} {#1} \cr {#2} \end{bmatrix}}

% Leave the next line alone. This is for my answer keys.
%\def\isanswerkey{1}

\input{../common/header474}
\if\isanswerkey1\input{macsse474-key}\fi

\begin{document}

\assignmentnametitlestuff


\if\isanswerkey0
{\bf Please follow the homework guidelines from A01.}

\fi


\begin{enumerate}

%\item (1.42) For languages $A$ and $B$ over the same alphabet $\Sigma$, let the {\bf shuffle} of $A$ and $B$ be the language
%\[\{w \mid w = a_1b_1\cdots a_kb_k, \textnormal{ where } a_1\cdots a_k \in A \textnormal{ and } b_1\cdots b_k \in B \textnormal{ for some $k \ge 0$, } \textnormal{ where each } a_i, b_i \in \Sigma^*\}.\]
%(Note: the difference from perfect shuffle is that in shuffle, $a_i, b_i \in \Sigma^*$, not $\Sigma$.)
%
%Show that the class of regular languages is closed under shuffle. 
%\solution{
%\if\isanswerkey1\solShuffle\fi
%}


\item (1.29c) Use the pumping lemma to show that this language is not regular:\\
$A_3 = \{{\tt a}^{2^n} \mid n \ge 0\}$ (Here, ${\tt a}^{2^n}$ means a string of $2^n$ {\tt a}'s.)
\solution{
Suppose $A_3$ is regular, let $p$ be the pumping length in the pumping lemma. Consider a string $s={\tt a}^{2^n}\in A_3$ where $2^n \ge p$. Note that $\mid s\mid=2^n\ge p$, so $s=xyz$ with the properties in pumping lemma. By the pumping lemma, $|xy|\le p$, let $\mid y\mid = k$ where $k > 0$, then string $|xy^4z|=2^n+3k$. Since $2^n+3k\ne 2^r$, which contradicts the pumping lemma, so that language $A_3$ is not regular.
}

\item (1.46ac) Prove that the following languages are not regular. You may use the pumping lemma and the closure of the class of regular languages under union, intersection, and complement.
\begin{enumerate}
\item $\{{\tt 0}^n{\tt 1}^m{\tt 0}^n \mid m,n \ge 0\}$
\solution{
Assume language $D=\{{\tt 0}^n{\tt 1}^m{\tt 0}^n \mid m,n \ge 0\}$ is regular. The let $p$ be the pumping length in the lemma. Consider a string $s={\tt 0}^p{\tt 1}{\tt 0}^p \in D$. Note that $|s|=2p+1 \ge p$ as in the properties in the pumping lemma. Since $|xy|\le p$, then $y\in {\tt 0}^+$. Thus, in the string $xy^2z$, there is more ${\tt 0}$s than ${\tt 1}$s, which contradicts the pumping lemma. Hence, the language $D$ is not regular. 
}
\item[(c)] $\{w \mid w \in \{0,1\}^* \textnormal{ is not a palindrome}\}$  \\
(A {\em palindrome} is a string that reads the same forward and backward.)
\solution{
Assume language $D=\{w \mid w \in \{0,1\}^* \textnormal{ is not a palindrome}\}$ is regular. Since language $A={\tt 0}^*{\tt 1}^*{\tt 0}^*$ is a regular language, then language $\overline{D} \cup A$ is also a regular language. The language of $\overline{D} \cup A$ is same as the language described in (a), which has been shown to be irregular, which contradicts the properties of the regular language. Hence, the language $D$ is not regular. 
}
\end{enumerate}

\item (1.70) We define the {\em avoids} operation for languages $A$ and $B$ to be
\[A \textnormal{ {\em avoids} } B = \{w \mid w \in A \textnormal{ and $w$ doesn't contain any string in $B$ as a substring}\}.\]
Prove that the class of regular languages is closed under the {\em avoids} operation.
\solution{
For two regular languages $A$ and $B$, assume they are all regular, so there must be a DFA\\ $M_A(Q_A, \Sigma, \delta_A, q_A, F_A)$ that recognizes language $A$ and a DFA $M_B(Q_B, \Sigma, \delta_B, q_B, F_B)$ that recognizes language $B$, where $A=L(M_A)$ and $B=L(M_B)$. So that we can construct a NFA that recongnizes $A ~avoids ~B$: $M(Q, \Sigma, \delta, q_0, F)$:
\begin{itemize}
    \item $Q$: For the DFA $M$, $Q=Q_A \cup Q_B \cup \left\{\varnothing \right\}$
    \item $\Sigma$: The alphabets are all same
    \item $\delta$: 
    \begin{equation*}
        \delta(q, d) = 
        \begin{cases}
            \delta(q, d) & \text{where } q \in Q_A \\
            \delta_A(q, d) \cup \left\{q_B\right\} & \text{where } q \in F_A, d=\epsilon \\
            \delta_B(q, d) & \text{where } q \in Q_B \\
            \varnothing & \text{where } q \in F_B
        \end{cases}
    \end{equation*}
    \item $q_0$: $q_0=q_A$
    \item $F$: $F=F_A$
\end{itemize}
Hence, there is a finite automata that recongnizes language $A ~avoids ~B$, which makes the language $A ~avoids ~B$ regular. So operation $avoids$ is closed. 
}

\item (1.71) 
\begin{enumerate}
\item Let $A = \{{\tt 0}^k u {\tt 0}^k \mid k \ge 1 \textnormal{ and } u \in \Sigma^*\}$. Show that $A$ is regular.
\solution{
Let language $A'={\tt 0}\Sigma^*{\tt 0}$, so that language $A=A'$. Since $A'$ is regular, then $A$ is also regular.
}
\item Let $B = \{{\tt 0}^k {\tt 1} u {\tt 0}^k \mid k \ge 1 \textnormal{ and } u \in \Sigma^*\}$. Show that $A$ is {\bf not} regular.
\solution{
Assume language $B$ is regular, let $p$ be the pumping length as in the pumping lemma. Consider a string $s={\tt 0}^p{\tt 1}\Sigma^*{\tt 0}^p$. Node that $|s|>p$, then $s=xyz$ as in pumping lemma. Since in the pumping lemma, $|xy|\le p$, then $y \in 0^+$. Then $xy^2z$ must have more ${\tt 0}$s, which contradicts the pumping lemma. Hence the language $B$ is not regular. 
}
\end{enumerate}



\end{enumerate}


\end{document}
