\documentclass[11pt]{article}

\usepackage{enumitem}


%%TO EDIT
\newcommand{\dueclassnumber}{17}
\newcommand{\assignmentnum}{7}

% CHANGE issolution{0} to issolution{1} for homework submission.
% WRITE solutions inside the \solution{} commands.
% or, you can use \if\issolution1 … \fi

\def\issolution{1}
\def\myname{Allen Liu} % My name goes here
\def\mysec{01} % Section number goes here
\def\myCM{374} % Campus Mailbox goes here

\input{../common/flags}
\newcommand\vv[2]{\begin{bmatrix} {#1} \cr {#2} \end{bmatrix}}

% Leave the next line alone. This is for my answer keys.
%\def\isanswerkey{1}

\input{../common/header474}
\if\isanswerkey1\input{macsse474-key}\fi


\usetikzlibrary{positioning}

\begin{document}

\assignmentnametitlestuff


\if\isanswerkey0
{\bf Please follow the homework guidelines from A01.}

\fi


\begin{enumerate}

\item (2.48) Let $\Sigma = \{0,1\}$. Let $C_1$ be the language of all strings that contain a {\tt 1} in their middle third. Let $C_2$ be the language of all strings that contain two {\tt 1}s in their middle third. In particular,
\begin{align*}
C_1 &= \{xyz \mid x,z \in \Sigma^* \textnormal{ and $y \in \Sigma^*{\tt 1}\Sigma^*$, where $|x| = |z| \ge |y|$}\} \cr
C_2 &= \{xyz \mid x,z \in \Sigma^* \textnormal{ and $y \in \Sigma^*{\tt 1}\Sigma^*{\tt 1}\Sigma^*$, where $|x| = |z| \ge |y|$}\}.
\end{align*}
\begin{enumerate}
\item Show that $C_1$ is a CFL.
\solution{
\if\isanswerkey1\solMiddleThirdHasOneCFL\fi
The language $C_1$ can be recognized by following CFG:
\begin{align*}
    S &\to S_1 \mid S_2 \\
    S_1 &\to XTX \mid T \\
    T &\to XXTX \mid {\tt 1}\\
    S_2 &\to XFX \mid F \\
    F &\to XFXX \mid {\tt 1} \\
    X &\to {\tt a} \mid  {\tt b} 
\end{align*}
}
\item Show that $C_2$ is not a CFL.
\solution{
\if\isanswerkey1\solMiddleThirdHasTwoOnesNotCFL\fi
Assume language $C_2$ is a context-free language, then pumping lemma applied. Let $p$ be the pumping length as in the pumping lemma. According to the lemma, string $s=uvxyz \in C_2$. Consider a string $s={\tt 0}^{p+1}{\tt 1}{\tt 0}^p{\tt 1}{\tt 0}^{p+1} \in C_2$, $|s|=3p+3 \ge p$ as in pumping lemma. Since $|vxy| \le p$, $v \in {\tt 0}^+$ and $y \in {\tt 0}^+$. So for a pumped string, let's say $s'=uv^2xy^2z$, at most 2 adjacent groups of {\tt 0}s would have been stretched. So this will result in the string has $|x|\not=|y|$ or $|y| > |z|$, which makes the $s' \not\in C_2$. Hence language $C_2$ is not regular.
}

\end{enumerate}

\item (3.1) Consider the Turing machine $M_2$, whose state diagram is below. (The notation $a \to b, \Rsym$ denotes reading $a$ and overwriting with $b$ at the current head location, then moving the head one step right.) In each case, give the sequence of configurations that $M_2$ enters when started on the indicated input string.

\begin{minipage}{.15\linewidth}
\begin{enumerate}
\item {\tt 0}
\item {\tt 00}
\item {\tt 000}
\item {\tt 000000}
\end{enumerate}

\end{minipage}\hfill
\begin{minipage}{.8\linewidth}
\begin{tikzpicture}[node distance=3cm]
\node [state, initial] (1) {$q_1$};
\node [state, right=3cm of 1] (2) {$q_2$};
\node [state, right=4cm of 2] (3) {$q_3$};
\node [state, below=2cm of 3] (4) {$q_4$};
\node [state, above right= 1.5cm and 2cm of 2] (5) {$q_5$};
\node [state, below=1cm of 2] (acc) {$\qacc$};
\node [state, below=2cm of 1] (rej) {$\qrej$};
\draw 
(1) edge[below] node{${\tt 0} \to \tvs,\Rsym$} (2)
(1) edge[left] node[align=center]{$\tvs \to \Rsym$\\${\tt x} \to \Rsym$} (rej)
(2) edge[below] node{${\tt 0} \to {\tt x},\Rsym$} (3)
(2) edge[right] node[align=center]{$\tvs \to \Rsym$} (acc)
(2) edge[loop above] node{${\tt x} \to \Rsym$} (2)
(3) edge[loop above] node{${\tt x} \to \Rsym$} (3)
(3) edge[bend right=10, left] node{${\tt 0} \to \Rsym$} (4)
(3) edge[above,sloped] node{$\tvs \to \Lsym$} (5)
(4) edge[bend right=10, right] node{${\tt 0} \to {\tt x}, \Rsym$} (3)
(4) edge[loop right] node{${\tt x} \to \Rsym$} (4)
(5) edge[above,sloped] node{$\tvs \to \Rsym$} (2)
(5) edge[loop above] node[align=center]{${\tt 0} \to \Lsym$ \\${\tt x} \to \Lsym$} (5)
(4) edge[bend left=7, above] node{$\tvs \to \Rsym$} (rej)
;
\end{tikzpicture}
\end{minipage}
\solution{
\if\isanswerkey1\solTMConfigurationSequence\fi
\begin{enumerate}
    \item \begin{align*}
        &q_1{\tt 0}\\
        &\tvs q_2\tvs\\
        &\qacc
    \end{align*}
    \item \begin{align*}
        &q_1{\tt 00}\\
        &\tvs q_2{\tt 0}\\
        &\tvs{\tt x}q_3\tvs\\
        &\qrej
    \end{align*}
    \item \begin{align*}
        &q_1{\tt 000}\\
        &\tvs q_2{\tt 00}\\
        &\tvs{\tt x}q_3{\tt 0}\\
        &\tvs{\tt x0}q_4\tvs\\
        &\qrej
    \end{align*}
    \item \begin{align*}
        &q_1{\tt 000000}\\
        &\tvs q_2{\tt 00000}\\
        &\tvs{\tt x}q_3{\tt 0000}\\
        &\tvs{\tt x0}q_4{\tt 000}\\
        &\tvs{\tt x0x}q_3{\tt 00}\\
        &\tvs{\tt x0x0}q_4{\tt 0}\\
        &\tvs{\tt x0x0x}q_3\tvs\\
        &\tvs{\tt x0x0}q_5{\tt x}\\
        &\tvs{\tt x0x}q_5{\tt 0x}\\
        &\tvs{\tt x0}q_5{\tt x0x}\\
        &\tvs{\tt x}q_5{\tt 0x0x}\\
        &\tvs q_5{\tt x0x0x}\\
        &q_5\tvs{\tt x0x0x}\\
        &\tvs q_2{\tt x0x0x}\\
        &\tvs{\tt x}q_2{\tt 0x0x}\\
        &\tvs{\tt xx}q_3{\tt x0x}\\
        &\tvs{\tt xxx}q_3{\tt 0x}\\
        &\tvs{\tt xxx0}q_4{\tt x}\\
        &\tvs{\tt xxx0x}q_4\tvs\\
        &\qrej
    \end{align*}
\end{enumerate}
}


\item (3.8) Give implementation-level descriptions (i.e., clear, complete English descriptions of the steps of the machine) of Turing machines that decide the following languages over the alphabet $\{{\tt 0},{\tt 1}\}$.
\begin{enumerate}
\item $\{w \mid w \textnormal{ contains an equal number of {\tt 0}s and {\tt 1}s}\}$
\item $\{w \mid w \textnormal{ contains twice as many {\tt 0}s as {\tt 1}s}\}$
\item $\{w \mid w \textnormal{ does not contain twice as many {\tt 0}s as {\tt 1}s}\}$
\end{enumerate}
\solution{
\if\isanswerkey1\solTMDescriptionsDecideLanguages\fi
\begin{enumerate}
    \item 
    \begin{enumerate}
        \item Scan to the right to make sure that the input has ${\tt 0} \ge 1$ and ${\tt 1} \ge 1$ or the input is empty. 
        \begin{itemize}
            \item If the input is empty $\to \qacc$
        \end{itemize}
        \item Scan to the right and cross of the first {\tt 0}
        \begin{itemize}
            \item If no {\tt 0} is found, scan to the right to find for {\tt 1}s
            \begin{itemize}
                \item If no {\tt 1} is found $\to \qacc$
                \item If a {\tt 1} has been found $\to \qrej$
            \end{itemize}
        \end{itemize}
        \item Rewind to the left and scan to the right to cross off {\tt 1}
        \begin{itemize}
            \item If no {\tt 1} has been found $\to \qrej$
        \end{itemize}
        \item Rewind to the left and repeat step ii. 
    \end{enumerate}
    \item
    \begin{enumerate}
        \item Scan to the right to make sure that the input has ${\tt 0} \ge 1$ and ${\tt 1} \ge 1$ or the input is empty. 
        \begin{itemize}
            \item If the input is empty $\to \qacc$
        \end{itemize}
        \item Scan to the right and cross of the first {\tt 1}
        \begin{itemize}
            \item If no {\tt 1} is found, scan to the right to find for {\tt 0}s
            \begin{itemize}
                \item If no {\tt 0} is found $\to \qacc$
                \item If a {\tt 0} has been found $\to \qrej$
            \end{itemize}
        \end{itemize}
        \item Rewind to the left and scan to the right to cross off first 2 {\tt 0}s
        \begin{itemize}
            \item If no {\tt 0} has been found $\to \qrej$
            \item If only one {\tt 0} has been found $\to \qrej$
        \end{itemize}
        \item Rewind to the left and repeat step ii. 
    \end{enumerate}
    \item
    \begin{enumerate}
        \item Scan to the right to make sure that the input has ${\tt 0} \ge 1$ and ${\tt 1} \ge 1$ or the input is empty. 
        \begin{itemize}
            \item If the input is empty $\to \qrej$
        \end{itemize}
        \item Scan to the right and cross of the first {\tt 1}
        \begin{itemize}
            \item If no {\tt 1} is found, scan to the right to find for {\tt 0}s
            \begin{itemize}
                \item If no {\tt 0} is found $\to \qrej$
                \item If a {\tt 0} has been found $\to \qacc$
            \end{itemize}
        \end{itemize}
        \item Rewind to the left and scan to the right to cross off first 2 {\tt 0}s
        \begin{itemize}
            \item If no {\tt 0} has been found $\to \qacc$
            \item If only one {\tt 0} has been found $\to \qacc$
        \end{itemize}
        \item Rewind to the left and repeat step ii. 
    \end{enumerate}
\end{enumerate}
}


\item (3.9) Let a $k$-PDA be a pushdown automata that has $k$ stacks. Thus a 0-PDA is an NFA and a 1-PDA is a conventional PDA. You already know that 1-PDAs are more powerful (recognize a larger class of languages) than 0-PDAs.
\begin{enumerate}
\item Show that 2-PDAs are more powerful than 1-PDAs.
\solution{
\if\isanswerkey1\solTwoPDAMorePowerfulOnePDA\fi
For a 2-PDAs, they can recognize some languages that is not recognizable on 1-PDA. For example, language $A=\left\{{\tt a}^n{\tt b}^n{\tt c}^n \mid n > 0\right\}$, it is not recognizable by 1-PDA since it is not CFL. It can be proved by the following procedure: \newline Assume language $A$ is a CFL, pumping lemma applied. Let $p$ be the pumping length as in pumping lemma. Then, $s=uvxyz \in A$ as in pumping lemma. Consider a string $s={\tt a}^p{\tt b}^p{\tt c}^p \in A$, $|s|=3p \ge p$. Since in the pumping lemma, $|vxy| \le p$, $v$ and $y$ at most contains 2 of the 3 kinds in $\left\{{\tt a}, {\tt b}, {\tt c}\right\}$. For a pumped string, say $s'=uv^2xy^2z$, since $|v| \ge 0$ and $|y| \ge 0$, it must have two kinds that have more symbols than the $3^{\textnormal{rd}}$ one. So $s' \not\in A$. Hence language $A$ is not a CFL. 
\newline For a 2-PDA, the language $A$ can be recognized by the following 2-PDA: 
\begin{itemize}
    \item Begin with pushing the start character \$ to both stacks
    \item When the input reads symbol {\tt a}, it pushed it into the both stacks
    \item When the input reads symbol {\tt b}, it popped a symbol from the first stack
    \item When the input reads symbol {\tt c}, it popped a symbol from the second stack
    \item When the input is empty and there is \$ in both stack $\to \qacc$
    \item Otherwise $\to \qrej$
\end{itemize}
Hence the 2-PDAs can recognize some languages that are not recognizable by 1-PDA, so it is more powerful than 1-PDA.
}
\item Show that 3-PDAs are not more powerful than 2-PDAs. You may use the result of Theorem 3.13 that any multi-tape Turing machine can be simulated by a single-tape Turing machine.\\
(Hint: Simulate a Turing machine tape with two stacks.) 
\solution{
\if\isanswerkey1\solThreePDANotMorePowerfulTwoPDA\fi
For a 2-PDA, at a given state $q_x$, where the first stack contains the characters $s_1$ and second stack contains the characters $s_2$. This is equivalent to a 2-tape Turing machine with first tape at $s_1q_x\tvs$ and second tape at $s_2q_x\tvs$. popping from a stack is equivalent to moving L and overwriting with character $\tvs$ and pushing ${\tt a}$ is equivalent with moving R and overwriting with character ${\tt a}$. So for a 3-PDA, it can be also converted into a 3-tape Turing machine. Since by Theorem 3.13, $k$-tape Turing machine can all be converted into a single-tape Turing machine. So both 2-PDAs and 3-PDAs can be converted into a single-tape Turing machine. Hence 3-PDAs is same powerful as 2-PDAs, so it is not more powerful than 2-PDAs.  
}
\end{enumerate}



\end{enumerate}





\end{document}
